The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-03 rockerverse.tex
only first mention uses \CRANpkg
preparing for submission, see #26

\textbf{\CRANpkg{stevedore}} is currently the most powerful Docker
client in R \citep{cran_stevedore}. It interfaces with the Docker daemon
over the Docker HTTP
API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}}
via a Unix socket on Linux or MacOS, over a named pipe on Windows, or
over an HTTP/TCP connection. The package is the only one not using
system calls to the \texttt{docker} CLI tool for managing images and
containers. The package thereby enables connections to remote Docker
instances without direct configuration of the local Docker daemon.
Furthermore using the API gives access to information in a structured
way, is system independent, and is likely more reliable than parsing
command line output. \CRANpkg{stevedore}'s own interface is
automatically generated based on the OpenAPI specification of the Docker
daemon, but it is still similar to the Docker CLI. The interface is
similar to R6 objects, in that an object of class
\code{"stevedore\_object"} has a number of functions attached to it that
can be called, and multiple specific versions of the Docker API can be
supported thanks to the automatic
generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

@@

\textbf{\pkg{containerit}}
(\url{https://github.com/o2r-project/containerit/}) attempts to take
this one step further and includes these tools to automatically create a
\texttt{Dockerfile} that can execute a given workflow
\citep{nust_containerit_2019}. \pkg{containerit} accepts an R object of
classes \code{"sessionInfo"} or \code{"session\_info"} as input and
provides helper functions to derive these from workflows, e.g., an R
script or R Markdown document, by analysing the session state at the end
of the workflow. It relies on the \pkg{sysreqs}
(\url{https://github.com/r-hub/sysreqs/}) package and it's mapping of
package system dependencies to platform specific installation package
names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \CRANpkg{stevedore} to streamline the user
interaction and improve the created \texttt{Dockerfile}s, e.g., by
running a container for the desired base image to extract the already
available R packages.

@@

\textbf{\CRANpkg{liftr}} aims to solve the problem of persistent
reproducible reporting in statistical computing based on the R Markdown
format \citep{xie2018} for dynamic documents
\citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for
collaborative documents and large-scale platforms for processing
documents. \CRANpkg{liftr} makes the document the main and sole workflow
control file and the only file that needs to be shared between
collaborators for consistent environments, e.g.~demonstrated in the
DockFlow project (\url{https://dockflow.org}). It introduces new fields
to the R Markdown document header, allowing users to manually declare
the dependencies, including versions, for rendering the document. The
package then generates a \texttt{Dockerfile} from this metadata and
provides a utitility function to render the document inside a Docker
container, i.e., \code{render\_docker("foo.Rmd")}. An RStudio addin even
allows compilation of documents with the single push of a button.

System dependencies are the domain of Docker, but for a full description
of the computing environment one must also manage the R version and the
the R packages. R versions are available via the versioned Rocker image
stack \citep{RJ-2017-065}.
\href{https://github.com/ColinFay/ronline}{r-online} leverages these
images and provides an app for helping users to detect breaking changes
between different R versions and for historic exploration of R. With a
standalone NodeJS app or
\href{https://srv.colinfay.me/r-online}{r-online}, the user can compare
a piece of code run in two separate versions of R. Internally, r-online
opens one or two Docker instances with the given version of R based on
Rocker images, executes a given piece of code, and returns the result to
the user. Regarding R package management, this can be achieved with
MRAN, or with packages such as \CRANpkg{checkpoint}
\citep{cran_checkpoint} and \CRANpkg{renv} \citep{renv2019}, which can
naturally be applied within images and containers. For example,
\CRANpkg{renv} helps users to manage the state of the R library in a
reproducible way, further providing isolation and portability. While
\CRANpkg{renv} does not cover system dependencies, the
\CRANpkg{renv}-based environment can be transferred into a container
either by restoring the environment based on the main configuration file
\texttt{renv.lock} or by storing the \CRANpkg{renv}-cache on the host
and not in the container \citep{ushey_using_2019}. With both the system
dependencies and R packages consciously managed in a Docker image, users
can start using containers as the \emph{only} environment for their
workflows, which allows them to work indepenently of physical
computers\footnote{Allowing them to be digital "nomads", cf. J.~Bryan's \href{https://github.com/jennybc/docker-why}{https://github.com/jennybc/docker-why}.}
and to assert a specific degree of confidence in the stability of a
developed software
\citep[cf. \texttt{README.Rmd} in][]{marwick_research_2017}.

@@

Second, the strong integration of \textbf{system libraries in core
packages} in the \href{https://www.r-spatial.org/}{R-spatial community}
makes containers essential for stable and proactive development of
common classes for geospatial data modelling and analysis. For example,
GDAL \citep{gdal_2019} is a crucial library in the geospatial domain.
GDAL is a system dependency allowing R packages such as \CRANpkg{sf},
which provides the core data model for geospatial vector data, or
\CRANpkg{rgdal}, to accommodate users to be able to read and write
hundreds of different spatial raster and vector formats
\citep{pebesma_simple_2018,cran_rgdal}. \CRANpkg{sf} and \CRANpkg{rgdal}
have hundreds of indirect reverse imports and dependencies and,
therefore, the maintainers spend a lot of effort trying not to break
them. Purpose-built Docker images are used to prepare for upcoming
releases of system libraries, individual bug reports, and for the lowest
supported versions of system
libraries\footnote{Cf. \href{https://github.com/r-spatial/sf/tree/master/inst/docker}{https://github.com/r-spatial/sf/tree/master/inst/docker}, \href{https://github.com/Nowosad/rspatial_proj6}{https://github.com/Nowosad/rspatial\_proj6}, and \href{https://github.com/r-spatial/sf/issues/1231}{https://github.com/r-spatial/sf/issues/1231}}.

@@

\CRANpkg{googleComputeEngineR} integrates with the R parallelisation
package \CRANpkg{future} \citep{cran_future} to enable this with only a
few lines or R

@@

actually needs to be re-executed. \CRANpkg{drake} has been demonstrated
to run inside containers for high

@@

Furthermore, \CRANpkg{drake} workflows have been shown to use
\CRANpkg{future} package's function \code{makeClusterPSOCK()} for
sending parts of the workflow to a Docker image for

@@

the \CRANpkg{future} package ensures that the host and worker can
connect for communicating over socket connections. \textbf{RStudio
Server Pro} includes a functionality called

@@

First, different packages use containers for the \textbf{deployment of R
and \CRANpkg{Shiny} apps}. \CRANpkg{Shiny} is a popular package for
creating interactive online dashboards with R, and it enables users with
very diverse backgrounds to create stable and user-friendly web
applications \citep{cran_shiny}. \emph{ShinyProxy}
(\url{https://www.shinyproxy.io/}) is an open source tool to deploy
Shiny apps in an enterprise context, where it features single sign-on,
but it can also be used in scientific use cases
\citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to
achieve scalability for multi-user scenarios with multiple apps.
ShinyProxy itself is written in Java to accommodate corporate
requirements and may itself run in a container for stability and
availability. The tool is built on ContainerProxy
(\url{https://www.containerproxy.io/}), which provides similar features
for executing long-running R jobs or interactive R sessions. The started
containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications
using Shinyproxy can be achieved, e.g., via
\href{https://docs.gitlab.com/ee/ci/pipelines.html}{GitLab pipelines}. A
pipeline can include building and checking R packages and Shiny apps.
After the code has passed the checks, Docker images are built and pushed
to GitLab's container registry. The pipeline finishes with triggering a
webhook on the server, where the deployment script is executed. The
script updates configurations, pulls the new Docker images, and restarts
the services.

@@

Third, containers can greatly facilitate the \textbf{deployment to
existing infrastructures}. \emph{Kubernetes}
(\url{https://kubernetes.io/}) is a container-orchestration system for
managing container-based application deployment and scaling. A
\emph{cluster} of containers, orchestrated as a single deployment, e.g.,
with Kubernetes, can mitigate limitations on request volumes or a
container occupied with computationally intensive task. A cluster
features load-balancing, autoscaling of containers across numerous
servers (in the cloud or on premise), and restarting failed ones. Many
organisations already use a Kubernetes cluster for other applications,
or a managed cluster can be acquired from service providers. Docker
containers are used within Kubernetes clusters to hold native code, for
which Kubernetes creates a framework around network connections and
scaling of resources up and down. Kubernetes can thereby host R
applications, big parallel tasks, or scheduled batch jobs in a scalable
way, and the deployment can even be triggered by changes to code
repositories \citep[i.e., CD, see][]{edmondson_r_2018}. The package
\pkg{googleKubernetesR}
(\url{https://github.com/RhysJackson/googleKubernetesR}) is a proof of
concept for wrapping the Google Kubernetes Engine API, Google's hosted
Kubernetes solution, in an easy-to-use R package. The package
\CRANpkg{analogsea} provides a way to programmatically create and
destroy cloud VMs on the \href{https://www.digitalocean.com/}{Digital
Ocean} platform \citep{analogsea_2019}. It also includes R wrapper
functions to install Docker in such a VM, manage images, and control
containers straight from R functions. These functions are translated to
Docker CLI commands and transferred transparently to the respective
remove machine using SSH. \CRANpkg{AzureContainers} is an umbrella
package that provides interfaces for three commercial services of
Microsoft's Azure Cloud, namely
\href{https://azure.microsoft.com/en-us/services/container-instances/}{Container
Instances} for running individual containers,
\href{https://azure.microsoft.com/en-us/services/container-registry/}{Container
Registry} for private image distribution, and
\href{https://azure.microsoft.com/en-us/services/kubernetes-service/}{Kubernetes
Service} for orchestrated deployments. While a package like
\CRANpkg{plumber} provides the infrastructure for turning an R workflow
into a web service, for production purposes it is usually necessary to
take into account scalability, reliability and ease of management.
AzureContainers provides an R-based interface to these features and,
thereby, simplifies complex infrastructure management to a number of R
function calls, given an Azure account with sufficient
credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.
\href{https://www.heroku.com/}{Heroku} is another cloud platform as a
service provider, and it supports container-based applications.
\texttt{heroku-docker-r}
(\url{https://github.com/virtualstaticvoid/heroku-docker-r}) is an
independent project providing a template for deploying R applications
based on Heroku's image stack, including multiple examples for
interfacing R with other programming languages. Yet the approach
requires a manual management of the computing environment.

Independent integrations of R for different cloud providers lead to
repeated efforts and code fragmentation. To mitigate these problems and
to avoid vendor lock-in motivated the
\href{https://www.openfaas.com/}{OpenFaaS project}. OpenFaas facilitates
the deployment of functions and microservices to Kubernetes or Docker
Swarm. It is language agnostic and provides auto-scaling, metrics, and
an API gateway. Reduced boilerplate code is achieved via templates.
Template for
R\footnote{See OpenFaaS R templates at \href{https://github.com/analythium/openfaas-rstats-templates}{https://github.com/analythium/openfaas-rstats-templates}.}
are provided based on Rocker's Debian and R-hub's
\href{https://github.com/r-hub/r-minimal}{r-minimal} Alpine images. The
templates use
\href{https://docs.docker.com/develop/develop-images/multistage-build/}{multi-stage
Docker builds} to combine R base images with the OpenFaaS `watchdog', a
tiny Golang web server. The watchdog marshals an HTTP request and
invokes the actual application. The R session uses \CRANpkg{plumber} or
similar packages for the API endpoint with packages and data preloaded,
thus minimizing response times.

@@

R has been historically viewed as a tool for analysis and scientific
research, but not for creating software that corporations can rely on
for production services. However, thanks to advancements in R running as
a web service, along with the ability to deploy R in Docker containers,
modern enterprises are now capable of having real-time machine learning
powered by R. A number of packages and projects have enabled R to
respond to client requests over TCP/IP and local socket servers, such as
\CRANpkg{Rserve} \citep{cran_rserve}, \CRANpkg{svSocket}
\citep{grosjean_sciviews_2019}, \href{http://www.rapache.net}{rApache}
and more recently \CRANpkg{plumber} (\url{https://www.rplumber.io/}) and
\pkg{RestRserve} (\url{http://restrserve.org}), which even processes
incoming requests in parallel with forked processes using
\CRANpkg{Rserve}. The latter two also provide documentation for
deployment with Docker or ready-to-use images with automated
builds\footnote{See \href{https://www.rplumber.io/docs/hosting.html\#docker}{https://www.rplumber.io/docs/hosting.html\#docker}, \href{https://hub.docker.com/r/trestletech/plumber/}{https://hub.docker.com/r/trestletech/plumber/} and \href{https://hub.docker.com/r/rexyai/restrserve/}{https://hub.docker.com/r/rexyai/restrserve/}.}.
These software allow other (remote) processes and programming languages
to interact with R and to expose R-based function in a service
architecture with HTTP APIs. APIs based on these packages can be
deployed with scalability and high availability using containers. This
pattern of deploying code matches those used by software engineering
services created in more established languages in the enterprise domain,
such as Java or Python, and R can be used alongside those languages as a
first-class member of a software engineering technical stack.

CARD.com implemented a web application for the optimisation of the
acquisition flow and the real-time analysis of debit card transactions.
The software used \CRANpkg{Rserve} and rApache and was deployed in
Docker containers. The R session behind \CRANpkg{Rserve} acted as a
read-only in-memory database, which was extremely fast and scalabale,
for the many concurrent rApache processes responding to the live-scoring
requests of various divisions of the company. Similarly deodorised R
scripts were responsible for the ETL processes and even the
client-facing email, text message and push notification alerts sent in
real-time based on card transactions. The related Docker images were
made available at \url{https://github.com/cardcorp/card-rocker}. The
images extended \texttt{rocker/r-base} and additionally entailed an SSH
client and a workaround for being able to mount SSH keys from the host,
Pandoc, the Amazon Web Services (AWS) SDK, and Java, which is required
by the AWS SDK. The AWS SDK allowed for running R consumers reading from
real-time data processing streams of
\href{https://aws.amazon.com/kinesis/}{AWS Kinesis}
\footnote{See useR!2017 talk \href{https://static.sched.com/hosted\_files/user2017/2f/AWR Kinesis at useR 2017.pdf}{"Stream processing with R in AWS"}.}.
The applications were deployed on Amazon Elastic Container Service
(\href{https://aws.amazon.com/ecs/}{ECS}). The main takeaways from using
R in Docker were not only that pinning the R package versions via MRAN
is important, but also that moving away from Debian testing to a
distribution with long-term support can be necessary. For the use case
at hand, this switch allowed for more control over upstream updates and
for minimising the risk of breaking the automated builds of the Docker
images and production jobs.

The AI @ T-Mobile team created a set of neural network machine learning
natural language processing models to help customer care agents manage
text-based messages for customers \citep{t-mobile_enterprise_2018}. For
example, one model quickly identifies whether a message is from a
customer
\citep[see \CRANpkg{Shiny}-based \href{https://secure.message.t-mobile.com/v1/shiny/is-customer/app/}{demo} further described by ][]{nolis_small_2019},
and others tell which customers are likely to make a repeat purchase. If
a data scientist creates a machine learning model and exposes it through
a \CRANpkg{plumber} API, then someone else on the marketing team could
write software that sends different emails depending on that real-time
prediction. The models are convolutional neural networks that use the
\CRANpkg{keras} package \citep{cran_keras} and run in a Rocker
container. The corresponding \texttt{Dockerfile}s are published
\href{https://github.com/tmobile/r-tensorflow-api}{on GitHub}. Since the
models power tools for agents and customers, they need to have extremely
high uptime and reliability. The AI @ T-Mobile team found that the
models performed well, and today these models power real-time services
that are called over a million times a day.

@@

Second, the idea of a common environment and partitioning allows for
using \emph{containers in teaching for secure execution and automated
testing} of submissions by students.
\href{https://dodona.ugent.be}{Dodona} is a web platform developed at
Ghent University that is used to teach students basic programming
skills, and it uses Docker containers to test submissions by students.
This means that both the code testing the students' submissions and the
submission itself are executed in a predictable environment, avoiding
compatibility issues between the wide variety of configurations used by
students. The containerisation is also used to shield the Dodona servers
from bad or even malicious code: memory, time and I/O~limits are used to
make sure students cannot overload the system. The web application
managing the containers communicates with them by sending configuration
information as a JSON document over standard input. Every Dodona Docker
image shares a \texttt{main.sh} file that passes through this
information to the actual testing framework, while setting up some error
handling. The testing process in the Docker containers sends back the
test results by writing a JSON document to its standard output channel.
In June 2019, R support was added to Dodona using an image derived from
the \texttt{rocker/r-base} image that sets up the \texttt{runner} user
and and \texttt{main.sh} file expected by
Dodona\footnote{\href{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}}.
It also installs the packages required for the testing framework and the
exercises so that this does not have to happen every time a student's
submission is evaluated. The actual testing of R exercises is done using
a custom framework loosely based on \CRANpkg{testthat}
\citep{wickham_testthat_2011}. During the development of the testing
framework, it was found that the \CRANpkg{testthat} framework did not
provide enough information to its reporter system to send back all the
fields required by Dodona to render its feedback. Right now, multiple
statistics courses are developing exercises to automate the feedback for
their lab classes.

>>>>

\textbf{\CRANpkg{stevedore}} is currently the most powerful Docker
client in R \citep{cran_stevedore}. It interfaces with the Docker daemon
over the Docker HTTP
API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}}
via a Unix socket on Linux or MacOS, over a named pipe on Windows, or
over an HTTP/TCP connection. The package is the only one not using
system calls to the \texttt{docker} CLI tool for managing images and
containers. The package thereby enables connections to remote Docker
instances without direct configuration of the local Docker daemon.
Furthermore using the API gives access to information in a structured
way, is system independent, and is likely more reliable than parsing
command line output. \pkg{stevedore}'s own interface is automatically
generated based on the OpenAPI specification of the Docker daemon, but
it is still similar to the Docker CLI. The interface is similar to R6
objects, in that an object of class \code{"stevedore\_object"} has a
number of functions attached to it that can be called, and multiple
specific versions of the Docker API can be supported thanks to the
automatic
generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

@@

\textbf{\pkg{containerit}}
(\url{https://github.com/o2r-project/containerit/}) attempts to take
this one step further and includes these tools to automatically create a
\texttt{Dockerfile} that can execute a given workflow
\citep{nust_containerit_2019}. \pkg{containerit} accepts an R object of
classes \code{"sessionInfo"} or \code{"session\_info"} as input and
provides helper functions to derive these from workflows, e.g., an R
script or R Markdown document, by analysing the session state at the end
of the workflow. It relies on the \pkg{sysreqs}
(\url{https://github.com/r-hub/sysreqs/}) package and it's mapping of
package system dependencies to platform specific installation package
names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \pkg{stevedore} to streamline the user
interaction and improve the created \texttt{Dockerfile}s, e.g., by
running a container for the desired base image to extract the already
available R packages.

@@

\textbf{\CRANpkg{liftr}} aims to solve the problem of persistent
reproducible reporting in statistical computing based on the R Markdown
format \citep{xie2018} for dynamic documents
\citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for
collaborative documents and large-scale platforms for processing
documents. \pkg{liftr} makes the document the main and sole workflow
control file and the only file that needs to be shared between
collaborators for consistent environments, e.g.~demonstrated in the
DockFlow project (\url{https://dockflow.org}). It introduces new fields
to the R Markdown document header, allowing users to manually declare
the dependencies, including versions, for rendering the document. The
package then generates a \texttt{Dockerfile} from this metadata and
provides a utitility function to render the document inside a Docker
container, i.e., \code{render\_docker("foo.Rmd")}. An RStudio addin even
allows compilation of documents with the single push of a button.

System dependencies are the domain of Docker, but for a full description
of the computing environment one must also manage the R version and the
the R packages. R versions are available via the versioned Rocker image
stack \citep{RJ-2017-065}.
\href{https://github.com/ColinFay/ronline}{r-online} leverages these
images and provides an app for helping users to detect breaking changes
between different R versions and for historic exploration of R. With a
standalone NodeJS app or
\href{https://srv.colinfay.me/r-online}{r-online}, the user can compare
a piece of code run in two separate versions of R. Internally, r-online
opens one or two Docker instances with the given version of R based on
Rocker images, executes a given piece of code, and returns the result to
the user. Regarding R package management, this can be achieved with
MRAN, or with packages such as \CRANpkg{checkpoint}
\citep{cran_checkpoint} and \CRANpkg{renv} \citep{renv2019}, which can
naturally be applied within images and containers. For example,
\pkg{renv} helps users to manage the state of the R library in a
reproducible way, further providing isolation and portability. While
\pkg{renv} does not cover system dependencies, the \pkg{renv}-based
environment can be transferred into a container either by restoring the
environment based on the main configuration file \texttt{renv.lock} or
by storing the \pkg{renv}-cache on the host and not in the container
\citep{ushey_using_2019}. With both the system dependencies and R
packages consciously managed in a Docker image, users can start using
containers as the \emph{only} environment for their workflows, which
allows them to work indepenently of physical
computers\footnote{Allowing them to be digital "nomads", cf. J.~Bryan's \href{https://github.com/jennybc/docker-why}{https://github.com/jennybc/docker-why}.}
and to assert a specific degree of confidence in the stability of a
developed software
\citep[cf. \texttt{README.Rmd} in][]{marwick_research_2017}.

@@

Second, the strong integration of \textbf{system libraries in core
packages} in the \href{https://www.r-spatial.org/}{R-spatial community}
makes containers essential for stable and proactive development of
common classes for geospatial data modelling and analysis. For example,
GDAL \citep{gdal_2019} is a crucial library in the geospatial domain.
GDAL is a system dependency allowing R packages such as \CRANpkg{sf},
which provides the core data model for geospatial vector data, or
\CRANpkg{rgdal}, to accommodate users to be able to read and write
hundreds of different spatial raster and vector formats
\citep{pebesma_simple_2018,cran_rgdal}. \pkg{sf} and \pkg{rgdal} have
hundreds of indirect reverse imports and dependencies and, therefore,
the maintainers spend a lot of effort trying not to break them.
Purpose-built Docker images are used to prepare for upcoming releases of
system libraries, individual bug reports, and for the lowest supported
versions of system
libraries\footnote{Cf. \href{https://github.com/r-spatial/sf/tree/master/inst/docker}{https://github.com/r-spatial/sf/tree/master/inst/docker}, \href{https://github.com/Nowosad/rspatial_proj6}{https://github.com/Nowosad/rspatial\_proj6}, and \href{https://github.com/r-spatial/sf/issues/1231}{https://github.com/r-spatial/sf/issues/1231}}.

@@

\pkg{googleComputeEngineR} integrates with the R parallelisation package
\CRANpkg{future} \citep{cran_future} to enable this with only a few
lines or R

@@

actually needs to be re-executed. \pkg{drake} has been demonstrated to
run inside containers for high

@@

Furthermore, \pkg{drake} workflows have been shown to use \pkg{future}
package's function \code{makeClusterPSOCK()} for sending parts of the
workflow to a Docker image for

@@

the \pkg{future} package ensures that the host and worker can connect
for communicating over socket connections. \textbf{RStudio Server Pro}
includes a functionality called

@@

First, different packages use containers for the \textbf{deployment of R
and \pkg{Shiny} apps}. \pkg{Shiny} is a popular package for creating
interactive online dashboards with R, and it enables users with very
diverse backgrounds to create stable and user-friendly web applications
\citep{cran_shiny}. \emph{ShinyProxy} (\url{https://www.shinyproxy.io/})
is an open source tool to deploy Shiny apps in an enterprise context,
where it features single sign-on, but it can also be used in scientific
use cases
\citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to
achieve scalability for multi-user scenarios with multiple apps.
ShinyProxy itself is written in Java to accommodate corporate
requirements and may itself run in a container for stability and
availability. The tool is built on ContainerProxy
(\url{https://www.containerproxy.io/}), which provides similar features
for executing long-running R jobs or interactive R sessions. The started
containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications
using Shinyproxy can be achieved, e.g., via
\href{https://docs.gitlab.com/ee/ci/pipelines.html}{GitLab pipelines}. A
pipeline can include building and checking R packages and Shiny apps.
After the code has passed the checks, Docker images are built and pushed
to GitLab's container registry. The pipeline finishes with triggering a
webhook on the server, where the deployment script is executed. The
script updates configurations, pulls the new Docker images, and restarts
the services.

@@

Third, containers can greatly facilitate the \textbf{deployment to
existing infrastructures}. \emph{Kubernetes}
(\url{https://kubernetes.io/}) is a container-orchestration system for
managing container-based application deployment and scaling. A
\emph{cluster} of containers, orchestrated as a single deployment, e.g.,
with Kubernetes, can mitigate limitations on request volumes or a
container occupied with computationally intensive task. A cluster
features load-balancing, autoscaling of containers across numerous
servers (in the cloud or on premise), and restarting failed ones. Many
organisations already use a Kubernetes cluster for other applications,
or a managed cluster can be acquired from service providers. Docker
containers are used within Kubernetes clusters to hold native code, for
which Kubernetes creates a framework around network connections and
scaling of resources up and down. Kubernetes can thereby host R
applications, big parallel tasks, or scheduled batch jobs in a scalable
way, and the deployment can even be triggered by changes to code
repositories \citep[i.e., CD, see][]{edmondson_r_2018}. The package
\pkg{googleKubernetesR}
(\url{https://github.com/RhysJackson/googleKubernetesR}) is a proof of
concept for wrapping the Google Kubernetes Engine API, Google's hosted
Kubernetes solution, in an easy-to-use R package. The package
\CRANpkg{analogsea} provides a way to programmatically create and
destroy cloud VMs on the \href{https://www.digitalocean.com/}{Digital
Ocean} platform \citep{analogsea_2019}. It also includes R wrapper
functions to install Docker in such a VM, manage images, and control
containers straight from R functions. These functions are translated to
Docker CLI commands and transferred transparently to the respective
remove machine using SSH. \pkg{AzureContainers} is an umbrella package
that provides interfaces for three commercial services of Microsoft's
Azure Cloud, namely
\href{https://azure.microsoft.com/en-us/services/container-instances/}{Container
Instances} for running individual containers,
\href{https://azure.microsoft.com/en-us/services/container-registry/}{Container
Registry} for private image distribution, and
\href{https://azure.microsoft.com/en-us/services/kubernetes-service/}{Kubernetes
Service} for orchestrated deployments. While a package like
\pkg{plumber} provides the infrastructure for turning an R workflow into
a web service, for production purposes it is usually necessary to take
into account scalability, reliability and ease of management.
AzureContainers provides an R-based interface to these features and,
thereby, simplifies complex infrastructure management to a number of R
function calls, given an Azure account with sufficient
credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.
\href{https://www.heroku.com/}{Heroku} is another cloud platform as a
service provider, and it supports container-based applications.
\texttt{heroku-docker-r}
(\url{https://github.com/virtualstaticvoid/heroku-docker-r}) is an
independent project providing a template for deploying R applications
based on Heroku's image stack, including multiple examples for
interfacing R with other programming languages. Yet the approach
requires a manual management of the computing environment.

Independent integrations of R for different cloud providers lead to
repeated efforts and code fragmentation. To mitigate these problems and
to avoid vendor lock-in motivated the
\href{https://www.openfaas.com/}{OpenFaaS project}. OpenFaas facilitates
the deployment of functions and microservices to Kubernetes or Docker
Swarm. It is language agnostic and provides auto-scaling, metrics, and
an API gateway. Reduced boilerplate code is achieved via templates.
Template for
R\footnote{See OpenFaaS R templates at \href{https://github.com/analythium/openfaas-rstats-templates}{https://github.com/analythium/openfaas-rstats-templates}.}
are provided based on Rocker's Debian and R-hub's
\href{https://github.com/r-hub/r-minimal}{r-minimal} Alpine images. The
templates use
\href{https://docs.docker.com/develop/develop-images/multistage-build/}{multi-stage
Docker builds} to combine R base images with the OpenFaaS `watchdog', a
tiny Golang web server. The watchdog marshals an HTTP request and
invokes the actual application. The R session uses \pkg{plumber} or
similar packages for the API endpoint with packages and data preloaded,
thus minimizing response times.

@@

R has been historically viewed as a tool for analysis and scientific
research, but not for creating software that corporations can rely on
for production services. However, thanks to advancements in R running as
a web service, along with the ability to deploy R in Docker containers,
modern enterprises are now capable of having real-time machine learning
powered by R. A number of packages and projects have enabled R to
respond to client requests over TCP/IP and local socket servers, such as
\CRANpkg{Rserve} \citep{cran_rserve}, \CRANpkg{svSocket}
\citep{grosjean_sciviews_2019}, \href{http://www.rapache.net}{rApache}
and more recently \pkg{plumber} (\url{https://www.rplumber.io/}) and
\pkg{RestRserve} (\url{http://restrserve.org}), which even processes
incoming requests in parallel with forked processes using
\CRANpkg{Rserve}. The latter two also provide documentation for
deployment with Docker or ready-to-use images with automated
builds\footnote{See \href{https://www.rplumber.io/docs/hosting.html\#docker}{https://www.rplumber.io/docs/hosting.html\#docker}, \href{https://hub.docker.com/r/trestletech/plumber/}{https://hub.docker.com/r/trestletech/plumber/} and \href{https://hub.docker.com/r/rexyai/restrserve/}{https://hub.docker.com/r/rexyai/restrserve/}.}.
These software allow other (remote) processes and programming languages
to interact with R and to expose R-based function in a service
architecture with HTTP APIs. APIs based on these packages can be
deployed with scalability and high availability using containers. This
pattern of deploying code matches those used by software engineering
services created in more established languages in the enterprise domain,
such as Java or Python, and R can be used alongside those languages as a
first-class member of a software engineering technical stack.

CARD.com implemented a web application for the optimisation of the
acquisition flow and the real-time analysis of debit card transactions.
The software used \pkg{Rserve} and rApache and was deployed in Docker
containers. The R session behind \pkg{Rserve} acted as a read-only
in-memory database, which was extremely fast and scalabale, for the many
concurrent rApache processes responding to the live-scoring requests of
various divisions of the company. Similarly deodorised R scripts were
responsible for the ETL processes and even the client-facing email, text
message and push notification alerts sent in real-time based on card
transactions. The related Docker images were made available at
\url{https://github.com/cardcorp/card-rocker}. The images extended
\texttt{rocker/r-base} and additionally entailed an SSH client and a
workaround for being able to mount SSH keys from the host, Pandoc, the
Amazon Web Services (AWS) SDK, and Java, which is required by the AWS
SDK. The AWS SDK allowed for running R consumers reading from real-time
data processing streams of \href{https://aws.amazon.com/kinesis/}{AWS
Kinesis}
\footnote{See useR!2017 talk \href{https://static.sched.com/hosted\_files/user2017/2f/AWR Kinesis at useR 2017.pdf}{"Stream processing with R in AWS"}.}.
The applications were deployed on Amazon Elastic Container Service
(\href{https://aws.amazon.com/ecs/}{ECS}). The main takeaways from using
R in Docker were not only that pinning the R package versions via MRAN
is important, but also that moving away from Debian testing to a
distribution with long-term support can be necessary. For the use case
at hand, this switch allowed for more control over upstream updates and
for minimising the risk of breaking the automated builds of the Docker
images and production jobs.

The AI @ T-Mobile team created a set of neural network machine learning
natural language processing models to help customer care agents manage
text-based messages for customers \citep{t-mobile_enterprise_2018}. For
example, one model quickly identifies whether a message is from a
customer
\citep[see \pkg{Shiny}-based \href{https://secure.message.t-mobile.com/v1/shiny/is-customer/app/}{demo} further described by ][]{nolis_small_2019},
and others tell which customers are likely to make a repeat purchase. If
a data scientist creates a machine learning model and exposes it through
a \pkg{plumber} API, then someone else on the marketing team could write
software that sends different emails depending on that real-time
prediction. The models are convolutional neural networks that use the
\CRANpkg{keras} package \citep{cran_keras} and run in a Rocker
container. The corresponding \texttt{Dockerfile}s are published
\href{https://github.com/tmobile/r-tensorflow-api}{on GitHub}. Since the
models power tools for agents and customers, they need to have extremely
high uptime and reliability. The AI @ T-Mobile team found that the
models performed well, and today these models power real-time services
that are called over a million times a day.

@@

Second, the idea of a common environment and partitioning allows for
using \emph{containers in teaching for secure execution and automated
testing} of submissions by students.
\href{https://dodona.ugent.be}{Dodona} is a web platform developed at
Ghent University that is used to teach students basic programming
skills, and it uses Docker containers to test submissions by students.
This means that both the code testing the students' submissions and the
submission itself are executed in a predictable environment, avoiding
compatibility issues between the wide variety of configurations used by
students. The containerisation is also used to shield the Dodona servers
from bad or even malicious code: memory, time and I/O~limits are used to
make sure students cannot overload the system. The web application
managing the containers communicates with them by sending configuration
information as a JSON document over standard input. Every Dodona Docker
image shares a \texttt{main.sh} file that passes through this
information to the actual testing framework, while setting up some error
handling. The testing process in the Docker containers sends back the
test results by writing a JSON document to its standard output channel.
In June 2019, R support was added to Dodona using an image derived from
the \texttt{rocker/r-base} image that sets up the \texttt{runner} user
and and \texttt{main.sh} file expected by
Dodona\footnote{\href{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}}.
It also installs the packages required for the testing framework and the
exercises so that this does not have to happen every time a student's
submission is evaluated. The actual testing of R exercises is done using
a custom framework loosely based on \CRANpkg{testthat}
\citep{wickham_testthat_2011}. During the development of the testing
framework, it was found that the \pkg{testthat} framework did not
provide enough information to its reporter system to send back all the
fields required by Dodona to render its feedback. Right now, multiple
statistics courses are developing exercises to automate the feedback for
their lab classes.
